---
title: "Interactive maps with Leaflet"
author: "Pankaj Shah"
date: "7/1/2019"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
# Below codes lets you set your R environment.

# file.edit(file.path("~", ".Rprofile")) # edit .Rprofile in HOME
# file.edit(".Rprofile") # edit project specific .Rprofile
```

Drawing interactive maps with Leaflet

```{r, message= FALSE, errorFALSE, warning = FALSE}
library(tidyverse)
library(leaflet)
library(stringr)
library(sf)
library(widgetframe)

options(digits = 3)
set.seed(1234)
theme_set(theme_minimal())

```

Leaflet is an open-source JavaScript library for creating interactive maps. Unlike static visualization packages such as ggplot2 or ggmap, Leaflet maps are fully interactive and can include features such as:

Interactive panning/zooming
Pop-up tooltips and labels
Highlighting/selecting regions
It is used by many news organizations and tech websites to visualize geographic data. The leaflet package for R enables the creation of interactive maps within R without learning how to write JavaScript code. The leaflet documentation is a handy walkthrough for the basics of creating Leaflet maps in R
```{r}
```
0.1 Basic usage
Leaflet maps are built using layers, similar to ggplot2.

Create a map widget by calling leaflet()
Add layers to the map using one or more of the layer functions (e.g. addTiles(), addMarkers(), addPolygons())
Repeat step 2 as many times as necessary to incorporate the necessary information
Display the map widget
A basic example is:

```{r}
m <- leaflet() %>%
  addTiles() %>%
  addMarkers(lng = -87.597241, lat = 41.789829,
             popup = "Saieh Hall of Economics")
m %>%
  frameWidget()
```

0.2 Basemaps
Like ggmap, leaflet supports basemaps using map tiles. By default, OpenStreetMap tiles are used.

```{r}
m <- leaflet() %>%
  setView(lng = -87.618994, lat = 41.875619, zoom = 12)
m %>%
  addTiles() %>%
  frameWidget()
```

0.3 Add markers
Markers are used to identify points on the map. Each point needs to be defined in terms of latitude/longitude coordinates. These can come from a variety of sources, most commonly either a map data file such as a shapefile or GeoJSON (imported using sf) or a data frame with latitude and longitude columns.

Let’s use the Chicago crimes data to draw a map of the city identifying the location of each reported homicide:

```{r}
library(dplyr)
chicago_crimes <- read.csv("https://raw.githubusercontent.com/shahnp/data/master/Chicago_crime_2017")
```


```{r}
glimpse(chicago_crimes)
```

```{r}
(homicides <- chicago_crimes%>%
  filter(`Primary.Type` == "HOMICIDE"))
```


```{r}
leaflet(data = homicides) %>%
  addTiles() %>%
  addMarkers() %>%
  frameWidget()
```
addMarkers() and related functions will automatically check data frames for columns called lng/long/longitude and lat/latitude (case-insensitively). If your coordinate columns have any other names, you need to explicitly identify them using the lng and lat arguments. Such as `addMarkers(lng = ~Longitude, lat = ~Latitude).

Without any customization, we get a basic map with each murder location indicated by a dropped pin. Each markers appearance can be customized, though the technical difficulty quickly ramps up. The awesome markers plugin offers the most straight-forward customizability options. Instead of using addMarkers(), use addAwesomeMarkers() to control the appearance of the markers using icons from the Font Awesome, Bootstrap Glyphicons, and Ion icons icon libraries. First you define the appearance of the icon using awesomeIcons(), then pass that as an argument to addAwesomeMarkers():

```{r}
icons <- awesomeIcons(
  icon = 'bolt',
  iconColor = 'orange',
  markerColor = "black",
  library = 'fa')

leaflet(data = homicides) %>%
  addTiles() %>%
  addAwesomeMarkers(icon = icons) %>%
  frameWidget()
```

One concern is that some neighborhoods have so many murders that the points overlap. One solution enabled by Leaflet’s interactivity is to cluster markers at varying levels of detail using the clusterOptions argument to addMarkers():

```{r}
leaflet(data = homicides) %>%
  addTiles() %>%
  addMarkers(clusterOptions = markerClusterOptions()) %>%
  frameWidget()
```

Alternatively, we could use circles using addCircleMarkers():

```{r}
leaflet(data = homicides) %>%
  addTiles() %>%
  addCircleMarkers() %>%
  frameWidget()
```

0.4 Add labels and popups
Each point can have text added to it using either a label (appears either on hover or statically) or a popup (appears only on click). For instance we can label each murder with the date/timestamp when it was originally reported.

```{r}
leaflet(data = homicides) %>%
  addTiles() %>%
  addMarkers(label = ~Date) %>%
  frameWidget()
```

If we only want the information to appear when we click on the point, we should instead use popup = ~Date:

```{r}
leaflet(data = homicides) %>%
  addTiles() %>%
  addMarkers(popup = ~Date) %>%
  frameWidget()
```

We can combine multiple pieces of information to create a custom popup message. Unfortunately this does require basic knowledge of writing HTML documents.

```{r}
homicides %>%
  mutate(popup = str_c(Date,
                       Block,
                       str_c("Location type:", `Location.Description`,
                             sep = " "),
                       sep = "<br/>")) %>%
  leaflet() %>%
  addTiles() %>%
  addMarkers(popup = ~popup) %>%
  frameWidget()
```
0.5 Add lines and shapes
Leaflet can also draw spatial lines and shapes from R and add them to maps. Given our previous exposure to sf and importing shapefiles using st_read(), let draw a map of Chicago with each community area outlined.




